A Survey of Range-Only SLAM Techniques and Sensors

### 2018-02-11 Paper.tex
Added Introduction

	\author{David Grabowsky, Samyak Shetty, Sam Shue, James M. Conrad}

@@

	\begin{abstract}

A common aspect in many autonomous robotics applications is addressing the problem of Simultaneous Localization and Mapping (SLAM). In most SLAM applications, the robot identifies distinguishable features in the environment, known as landmarks, within the environment through sensors such as cameras and laser rangefinders. Sensing these landmarks provide the robot with information such as range and bearing, which are used in the estimation proses of the robot's pose and the landmark locations within the map. However, some applications use landmarks which are only capable of providing range information, such as wireless and acoutstic beacons. In this survey, the variety of techniques used to solve the SLAM problem with respect to range only sensors (RO-SLAM) is explored.

	\end{abstract}

@@

	In outdoor SLAM applications, such as self-driving cars and autonomous farming equipment, global positioning satellite (GPS) readings are used in addition to the vehicle's odometry. While GPS has low resolution, it proves useful as it provides location information, which can help compensate for the accumulated error from drift in odometric sensors. However, GPS isn't available in all environments, such as indoors and underwater. Instead, wireless signals and sonar beacons are often used in indoors and underwater environments, respectively. While these beacons provide sensor information to the robotic vehicle, they don't inherently provide location information unless their positions are accurately known beforehand and are certain to never change. Wireless and sonar beacons are capable of providing range measurements instead.

	\section{range only sensors/tech}

	One of the key features of the systematic localization and mapping problem are landmarks. In RO SLAM the only information that can be obtained from landmarks is range data. This range data must then be interpreted by the robot in some way that will be advantageous towards the localization and mapping process. Beacons are a popular implementation of landmarks. Beacons refer to any devices which involves a signal transmission that can be used for the localization and mapping process without line of sight. The most popular beacons can largely be broken down into two categories: acoustic based and radio frequency based.

@@

	Acoustic ranging systems are primary used with robotics and beacons that operate in underwater locations. The reason for this that radio waves experience strong attenuations, especially in salt water environments. As \cite{Partan2007} states, the issue of attenuations makes radio waves extremely impractical, thus making acoustics the logical alternative. Acoustic still suffer from a range of problems such as multipathing and propagation delays, but are one of the only meaningful alternatives. Papers such as those by \cite{Erol-Kantarci2011} and \cite{Akyildiz2005} explore a wide variety of issue and possible techniques that relate to acoustic sensors, such as the added complexity of assuming a 2D environment vs a 3D environment, or design challenge that accompany underwater deployment like spatial correlation and power.

	An interesting beacon for underwater localization that is not specifically range only or acoustic based is the Dive'N'Rise (DNR) beacons presented by \cite{Erol2007}. These DNR beacons rise to the surface, receive their location through GPS, then sink into the water and transmit their coordinates. The distance to an agent and the transmitting beacon can be determined via time of flight from the transmissions. A localization implementation using these beacons was accomplished by \cite{Erol2008}. Again while this is not purely range only or acoustic based it was found to be interesting solution worth mentioning.

%	\includegraphics[width=90mm]{DNR.png}

%	\includegraphics[width=90mm]{DNR.png}

	\caption{DNR being used to implement Dive and Rise Localization  \cite{Erol-Kantarci2011}}

	\label{DNR}

	\subsection{Radio Frequency}
	Radio frequency encompass a massive spectrum of devices capable of transmitting or receiving transmission on the radio frequency. There are two methods primarily utilized to determine distance from radio frequency. One is by examining the received signal strength indicator (RSSI) and the other is the by measuring the time of flight of the signal. A system presented by \cite{Padmanabhan2000} processes rf signal strength and utilizes nearest neighbors heuristics to achieve localization based on said signal strength.  \cite{Kurth2003a} describes a method for using very small RF tags as beacons throughout the environment.  As a robot moves throughout an environment it emits a query to which any RF tags that are within the transmission range respond to. The robot then estimates the distance to the tag based on the delay. The response from the RF tags include a data tag which contains a unique ID number.

	Wireless sensors networks (WSN) are also a very popular choice, especially since many of them come with the ability to transmit RSSI. \cite{Baronti2007}, \cite{Khanafer2014}, \cite{Akyildiz2002}, and \cite{Xia2011} each presents surveys relating to WSNs.  \cite{Baronti2007} and \cite{Akyildiz2002} both provide very well developed survey that discuss hardware, protocols, standards, security, energy efficiency and much more. \cite{Baronti2007} specifically focuses on state of the art 802.15.4 and ZigBee standards. \cite{Xia2011} examines the limitations of original IEEE 802.15.4 MAC and reviews adaptive methods that have been used to overcome these limitations. \cite{Khanafer2014} specifically survey methods meant to improve the performance of IEEE 802.15.4 MAC protocol.

@@

\section{Beacon Initialization}
Beacons initialization refers to the precursor process to the EKF where beacons (or landmarks) whose initial positions are not known are estimated. One of the complications with RO-SLAM involves the issue of the EKF, which is used by many implementations, being  sensitive to poorly initialized landmarks as they are used as the basis for the convergence of an accurate estimate. Poorly approximated landmarks results in the divergence of the filter. These methods are further  divided into delayed and undelayed initializations.
\subsection{Delayed/Offline initializations}
These are methods that perform delayed localizations  where initially the robot moves around offline collects some measurements and then initializes the location of the landmarks. This can be undersirable as it causes a delay in the localization of the robot whenever there is a immediate real time application such as in autonomous vehicles. Localization can be performed only once a accurate estimations of the beacons are made. There are several offline algorithms that perform localization such as the least squares approach in \cite{Newman},a posed based ekf as mentioned in \cite{Webster} , a moving horizon estimation as mentioned in \cite{SenWang2013} etc. The most popular ones are mentioned below.

\begin{figure}
	\centering
	\includegraphics[height=40mm,width=90mm]{Trilateration_methods.JPG}
	\caption{The red arrow represents the robot.The 'X' represents the origin of the robot. The blue line shows the traversed path.The star represents the landmark. }

\end{figure}
\subsubsection{Trilateration}
\cite{Thrun2002} mentions a basic trilateration technique involving geometric intersection. Rings are drawn at different points of the robots path as shown in Fig[a] , where the radius of the rings depend on the strength of the received signals from the beacons. This method of initialization is highly ambiguous as the rings can have more than 1 point of intersecting as shown in Fig[b]  or no  intersecting points as shown in Fig[c]. %more paperes needed
\subsubsection{Particle filter}
This takes a more probabilistic approach to estimating the initial location of the beacons. As mentioned in\cite{Thrun2002a} initially the entire space is represented by particles. These particles are resampled based on importance weights every iteration. There are several resampling methods mentioned in \cite{Li2015}. Eventually the particle set converges to the accurate location of the beacons. Though this method of localization can be computationally expensive it is preferred whenever the noise involved is non Gaussian.

@@

	\section{Methods and Their Implementations}

@@

	\section{Conclusion}

>>>>

	\author{\IEEEauthorblockN{David Grabowsky\IEEEauthorrefmark{1}, Samyak Shetty\IEEEauthorrefmark{2}, Sam Shue\IEEEauthorrefmark{3}, James M. Conrad}
		\IEEEauthorblockA{Department of Electrical and Computer Engineering\\
			University of North Carolina at Charlotte\\
			Charlotte, NC, USA\\
			\IEEEauthorrefmark{1}dgrabows@uncc.edu, \IEEEauthorrefmark{2}nshetty@uncc.edu, \IEEEauthorrefmark{3} slshue@uncc.edu, \IEEEauthorrefmark{4}jmconrad@uncc.edu}
	}

@@

	\begin{abstract}		
		A common aspect in many autonomous robotics applications is addressing the problem of Simultaneous Localization and Mapping (SLAM). In most SLAM applications, the robot identifies distinguishable features in the environment, known as landmarks, within the environment through sensors such as cameras and laser rangefinders. Sensing these landmarks provide the robot with information such as range and bearing, which are used in the estimation proses of the robot's pose and the landmark locations within the map. However, some applications use landmarks which are only capable of providing range information, such as wireless and acoutstic beacons. In this survey, the variety of techniques used to solve the SLAM problem with respect to range only sensors (RO-SLAM) is explored.
	\end{abstract}

@@

		In outdoor SLAM applications, such as self-driving cars and autonomous farming equipment, global positioning satellite (GPS) readings are used in addition to the vehicle's odometry. While GPS has low resolution, it proves useful as it provides location information, which can help compensate for the accumulated error from drift in odometric sensors. However, GPS isn't available in all environments, such as indoors and underwater. Instead, wireless signals and sonar beacons are often used in indoors and underwater environments, respectively. While these beacons provide distance estimations to the robotic vehicle, they do not inherently provide location information as with GPS. However, these distances are often still useful as they can be used to compensate for the drift that occurs from measuring a vehicle's displacement through integration of inertial sensor readings. Most SLAM algorithms utilize these range values as a corrective measurement, the same way traditional algorithms use sensed landmarks. However, rather than using range and orientation to landmarks, beacons are only capable of providing range. SLAM algorithms which utilize these beacons are referred to as Range-Only SLAM (RO-SLAM).

		While the loss of orientation from traditional, vision based landmarks prevents refinement of the robot's orientation, range only beacons solve two major problems encountered in traditional landmark-based approaches: landmark association and landmark scaling. The landmark association problem occurs when a detected landmark has similar features to previously recorded landmark and is mistaken for the other, generating huge errors in the estimation of both the robot and the landmark. This problem is avoided using beacons, as they have some kind of network address which is broadcast during transmissions, so there is no way to confuse which range is associated with which landmark. The scaling problem occurs when a SLAM algorithm runs for a long time and more and more landmarks are appended to the map. As more landmarks are added, the amount of memory needed to store the positions and distributions of these landmarks increases, as well as computation time. When using beacons, a finite number of landmarks exists, and the total size of the map is known before deployment. In addition, fewer landmarks are needed in beacon-based applications, as most beacons do not require line-of-sight to be measured, unlike visual landmarks. Since the robot can observe the landmarks more often, it doesn't need to add nearly as many to ensure accurate position estimation.

		This paper surveys the most popular techniques in RO-SLAM applications. Before discussing RO-SLAM techniques, the different types of range sensors are introduced in Section \ref{range only sensors}, followed by initialization of these beacons in Section \ref{beacon initialization}. Solutions to the RO-SLAM state estimation problem are covered in Section \ref{methods and their implementations}. Finally, the conclusion in Section \ref{conclusion} closes with final thoughts regarding the state of the field of RO-SLAM.

	\section{range only sensors}
	\label{range only sensors}
		There are a variety of range sensors which can be utilized as beacons for RO-SLAM. Beacons refer to any device which involves a signal transmission that can be used for the range estimation without line of sight. The most popular beacons can largely be broken down into two categories: acoustic based and radio frequency based. These beacons serve as landmarks within larger SLAM algorithms. The following section surveys the different types of sensors which can be utilized as range beacons.

@@

			Acoustic ranging systems are primarily used with autonomous underwater vehicles (AUV) and sonar beacons that operate in underwater locations. The reason for this that radio waves experience strong attenuations, especially in salt water environments. As \cite{Partan2007} states, the issue of attenuations makes radio waves extremely impractical, thus making acoustics the logical alternative. Acoustic still suffer from a range of problems such as multipathing and propagation delays, but are one of the only meaningful alternatives. Papers such as those by \cite{Erol-Kantarci2011} and \cite{Akyildiz2005} explore a wide variety of issues and possible solutions that relate to acoustic sensors, such as the added complexity of assuming a 2D environment versus a 3D environment, or the design challenges that accompany underwater deployment with regards to spatial correlation and power consumption.

			A unique implementation of an underwater acoustic beacon is the Dive'N'Rise (DNR) beacon presented by \cite{Erol2007}. These DNR beacons rise to the surface, receive their location through GPS, then sink into the water and transmit their coordinates. The distance to an agent and the transmitting beacon can be determined via time of flight from the transmissions. A localization implementation using these beacons was accomplished by \cite{Erol2008}.

			  	\begin{figure}[h!]
					\centering
				%	\includegraphics[width=90mm]{DNR.png}
				%	\includegraphics[width=90mm]{DNR.png}	
					\caption{DNR being used to implement Dive and Rise Localization  \cite{Erol-Kantarci2011}}
					\label{DNR}
				\end{figure}

		\subsection{Radio Frequency}
			Radio frequency encompass a massive spectrum of devices capable of transmitting or receiving transmission on the radio frequency. There are two methods primarily utilized to determine distance from radio frequency. One is by examining the received signal strength indicator (RSSI) and the other is the by measuring the time of flight of the signal. A system presented by \cite{Padmanabhan2000} processes RF signal strength and utilizes nearest neighbors heuristics to achieve localization based on said signal strength.  \cite{Kurth2003a} describes a method for using very small RF tags as beacons throughout the environment.  As a mobile robot moves throughout an environment it emits a query to which any RF tags that are within the transmission range respond. The robot then estimates the distance to the tag based on the delay. The response from the RF tags include a data tag which contains a unique ID number.

			Wireless sensors networks (WSN) are also a very popular choice, especially since many of them come with the ability to transmit RSSI. \cite{Baronti2007}, \cite{Khanafer2014}, \cite{Akyildiz2002}, and \cite{Xia2011} each present surveys relating to WSNs.  \cite{Baronti2007} and \cite{Akyildiz2002} both provide very well developed surveys that discuss hardware, protocols, standards, security, energy efficiency and much more. \cite{Baronti2007} specifically focuses on state of the art 802.15.4 and ZigBee standards. \cite{Xia2011} examines the limitations of original IEEE 802.15.4 MAC and reviews adaptive methods that have been used to overcome these limitations. \cite{Khanafer2014} specifically survey methods meant to improve the performance of IEEE 802.15.4 MAC protocol.

@@

	\section{Beacon Initialization}
		\label{beacon initialization}
		The core motivation of SLAM algorithms is no a priori knowledge of the operating environment. This includes locations of landmarks, including beacons. Because of this, landmarks are appended to the map as they are discovered in traditional SLAM algorithms, using the robot's estimated pose and the sensed distance and orientation to estimate an initial position of the landmark. The landmark's positions is then refined as the algorithm continues to run.

		The same process occurs in RO-SLAM, however, the landmark's initial position is not so easily determined, as the range-only beacon has an unknown orientation relative to the robot. Therefore, the landmark could reside approximately anywhere within the sensed distance radius. To compensate for this, most beacons go through an initialization phase before being appended to the map and utilized in the RO-SLAM algorithm. This initialization phase is usually performed over several iterations until the position converges to an acceptable degree.

		One RO-SLAM complication in particular involves the issue of EKF-SLAM, which is used by many implementations, being  sensitive to poorly initialized landmarks as they are used as the basis for the convergence of an accurate estimate. Poorly approximated landmarks results in the divergence of the filter.

		Presented here are the most effective and commonly used methods to initialize range beacons. These methods are further divided into delayed and undelayed initializations.

		\subsection{Delayed/Offline initializations}
			These are methods that perform delayed localizations  where initially the robot moves around offline collects some measurements and then initializes the location of the landmarks. This can be undersirable as it causes a delay in the localization of the robot whenever there is a immediate real time application such as in autonomous vehicles. Localization can be performed only once a accurate estimations of the beacons are made. There are several offline algorithms that perform localization such as the least squares approach in \cite{Newman},a posed based ekf as mentioned in \cite{Webster}, a moving horizon estimation as mentioned in \cite{SenWang2013} etc. The most popular ones are mentioned below.

		\begin{figure}
			\centering
			\includegraphics[height=40mm,width=90mm]{Trilateration_methods.JPG}
			\caption{The red arrow represents the robot.The 'X' represents the origin of the robot. The blue line shows the traversed path.The star represents the landmark. }
		\end{figure}

	\subsubsection{Trilateration}
		\cite{Thrun2002} mentions a basic trilateration technique involving geometric intersection. Rings are drawn at different points of the robots path as shown in Fig[a] , where the radius of the rings depend on the strength of the received signals from the beacons. This method of initialization is highly ambiguous as the rings can have more than 1 point of intersecting as shown in Fig[b]  or no  intersecting points as shown in Fig[c]. %more papers needed

	\subsubsection{Particle filter}
		This takes a more probabilistic approach to estimating the initial location of the beacons. As mentioned in\cite{Thrun2002a} initially the entire space is represented by particles. These particles are resampled based on importance weights every iteration. There are several resampling methods mentioned in \cite{Li2015}. Eventually the particle set converges to the accurate location of the beacons. Though this method of localization can be computationally expensive it is preferred whenever the noise involved is non Gaussian.

@@

	\section{Methods and Their Implementations}
	\label{methods and their implementations}

@@

	\section{Conclusion}
	\label{conclusion}
